R Notebook

### 2016-08-05 finalwriteup/draft.rmd
can't remember what I changed -- some grammar

Engaging in loss making jobs for fixed fees is a major problem in consulting, particularly in the competitive construction industry. This thesis investigates whether machine learning techniques applied to a company's passively collected internal data could help avoid loss making jobs tactfully choose when to enforce stricter contracts. It was found that in a specific decision framework, a case study's profits could be improved 9% by avoiding approximately 4% of projects. Alternative decision frameworks are also proposed and evaluated. The internal data collected by the case study company described each of their projects in terms of employee positions, fraction of project time completed by each employee, client characteristics, project characteristics (time span, hours, technical details) and invoicing history. Algorithmic methods such as Logistic Regression, Random Forests, Boosted Trees, Naive Bayes, and Bayesian Networks were applied as well as blended combinations of these methods. A decision scenario which" rejected projects above a sequence of tested thresholds was run in order to find the optimal threshold for profit improvements. This process was repeated on hundreds of models built from different training subsets of the data in order to obtain a 95% statistical confidence interval of profit improvements. The blended Logistic Regression model outperformed other methods and produced a 95% confidence interval of 6.5 - 11.5% profit improvements. Further work is recommended to test user input estimation and user interface designs. The findings from this research have the potential to assist managers in reducing losses by highlighting risky projects and guiding project-based changes to fee structures.

@@

ggplot(reduc.totAUC, aes(x= factor(L1), y= value)) + 
        geom_violin(aes(fill = sigdiff, colour = sigdiff), alpha = 0.8) + 
        stat_summary(fun.y="mean", geom="point", colour = "gray30", shape =3) +
        stat_summary(aes(label = round(..y.., 3)), fun.y="mean",
                     geom="text", colour = "gray20", size =3, 
                     hjust = -.2, angle = 40, vjust = -.2
                     ) +
        scale_fill_manual(values = c('sienna1','navajowhite'),
                          name = "Significantly different from boosted trees") +
        scale_colour_manual(values = c('sienna1','navajowhite'),
                          name = "Significantly Different from Boosted Trees") +
        labs(x = "Method", y = "Area Under Curve (AUC)",
             title = "AUC Distribution from 100 models of Each Method") +
        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
                         "log" = "Logistic Regression",
                         "naiive" = "Naive Bayes", "rf" = "Random Forest")) +
        theme(legend.position = "bottom")

@@

mice.totAUC <- readRDS("report_data/mice_df_auc.rds")

@@

cfor.varimp = readRDS("report_data/cfor_blend_varimp.rds")
cfor.varimp <- mutate(cfor.varimp, var = revalue(var, c("norm.log" = "log.pred",
                                         "norm.rf" = "rf.pred",
                                         "norm.boost" = "boost.pred")))

@@

log.varimp = readRDS("report_data/logblend.rds")
log.varimp = summary(log.varimp)$coefficients %>% as.data.frame
colnames(log.varimp)[names(log.varimp) %in% 'Pr(>|z|)']<-'P_value'
log.varimp = add_rownames(log.varimp)
log.varimp = log.varimp %>% filter(P_value <= 0.05) %>% 
        select(rowname, Estimate, P_value) %>% arrange(P_value)

@@

rf.varimp = readRDS("report_data/rfblend.rds")
rf.varimp = importance(rf.varimp, type = 1) %>% as.data.frame
rf.varimp = add_rownames(rf.varimp)
rf.varimp = arrange(rf.varimp, desc(MeanDecreaseAccuracy))
rf.varimp <- mutate(rf.varimp, rowname = revalue(rowname, c("norm.log" = "log.pred",
                                         "norm.rf" = "rf.pred",
                                         "norm.boost" = "boost.pred")))
# order levels
rf.varimp$rowname <- factor(rf.varimp$rowname,
                               levels= rf.varimp$rowname)

@@

boost.varimp = readRDS("report_data/boostblend.rds")
boost.varimp = add_rownames(boost.varimp)
boost.varimp <- mutate(boost.varimp, rowname = revalue(rowname, 
                                                       c("norm.log" = "log.pred",
                                         "norm.rf" = "rf.pred",
                                         "norm.boost" = "boost.pred")))
# order levels
boost.varimp$rowname <- factor(boost.varimp$rowname,
                               levels= boost.varimp$rowname)

@@

auc.list <- readRDS("report_data/blend_auc2.rds")
means = lapply(auc.list, mean) %>% unlist
greater = means[means> means["orig.log"]] %>% names
auc.melt = melt(auc.list)
auc.melt$greater = ifelse(auc.melt$L1 %in% greater, TRUE, FALSE)

@@

```{r blend_vs_initial}
auc.150 = readRDS("report_data/blend_auc150.rds")
mice.totAUC <- readRDS("report_data/mice_df_auc.rds")
compare.blends <- append(auc.150, mice.totAUC)

@@

reject.plot <- readRDS('report_data/reject_plot.rds')

@@

conf.plot <- readRDS("report_data/conf_plot.rds")

>>>>

Engaging in loss making jobs for fixed fees is a major problem in consulting, particularly in the competitive construction industry. This thesis investigates whether machine learning techniques applied to a company's passively collected internal data could help avoid loss making jobs or help tactfully choose when to enforce stricter contracts. It was found that in a specific decision framework, a case study's profits could be improved 9% by avoiding approximately 4% of projects. Alternative decision frameworks are also proposed and evaluated. The internal data collected by the case study company described each of their projects in terms of employee positions, fraction of project time completed by each employee, client characteristics, project characteristics (time span, hours, technical details) and invoicing history. Algorithmic methods such as Logistic Regression, Random Forests, Boosted Trees, Naive Bayes, and Bayesian Networks were applied as well as blended combinations of these methods. A decision scenario which" rejected projects above a sequence of tested thresholds was run in order to find the optimal threshold for profit improvements. This process was repeated on hundreds of models built from different training subsets of the data in order to obtain a 95% statistical confidence interval of profit improvements. The blended Logistic Regression model outperformed other methods and produced a 95% confidence interval of 6.5 - 11.5% profit improvements. Further work is recommended to test user input estimation and user interface designs. The findings from this research have the potential to assist managers in reducing losses by highlighting risky projects and guiding project-based changes to fee structures.

@@

ggplot(reduc.totAUC, aes(x= factor(L1), y= value)) + 
        geom_violin(aes(fill = sigdiff, colour = sigdiff), alpha = 0.8) + 
        stat_summary(fun.y="mean", geom="point", colour = "gray30", shape =3) +
        stat_summary(aes(label = round(..y.., 3)), fun.y="mean",
                     geom="text", colour = "gray20", size =3, 
                     hjust = -.2, angle = 40, vjust = -.2
                     ) +
        scale_fill_manual(values = c('sienna1','navajowhite'),
                          name = "Significantly different from Boosted Trees") +
        scale_colour_manual(values = c('sienna1','navajowhite'),
                          name = "Significantly different from Boosted Trees") +
        labs(x = "Method", y = "Area Under Curve (AUC)",
             title = "AUC Distribution from 100 models of Each Method") +
        scale_x_discrete(labels = c("boost" = "Boosted Trees", 
                         "log" = "Logistic Regression",
                         "naiive" = "Naive Bayes", "rf" = "Random Forest")) +
        theme(legend.position = "bottom")

@@

mice.totAUC <- readRDS("mice_df_auc.rds")

@@

cfor.varimp = readRDS("cfor_blend_varimp.rds")
cfor.varimp <- mutate(cfor.varimp, var = revalue(var, c("norm.log" = "log.pred",
                                         "norm.rf" = "rf.pred",
                                         "norm.boost" = "boost.pred")))

@@

log.varimp = readRDS("logblend.rds")
log.varimp = summary(log.varimp)$coefficients %>% as.data.frame
colnames(log.varimp)[names(log.varimp) %in% 'Pr(>|z|)']<-'P_value'
log.varimp = add_rownames(log.varimp)
log.varimp = log.varimp %>% filter(P_value <= 0.05) %>% 
        select(rowname, Estimate, P_value) %>% arrange(P_value)

@@

rf.varimp = readRDS("rfblend.rds")
rf.varimp = importance(rf.varimp, type = 1) %>% as.data.frame
rf.varimp = add_rownames(rf.varimp)
rf.varimp = arrange(rf.varimp, desc(MeanDecreaseAccuracy))
rf.varimp <- mutate(rf.varimp, rowname = revalue(rowname, c("norm.log" = "log.pred",
                                         "norm.rf" = "rf.pred",
                                         "norm.boost" = "boost.pred")))
# order levels
rf.varimp$rowname <- factor(rf.varimp$rowname,
                               levels= rf.varimp$rowname)

@@

boost.varimp = readRDS("boostblend.rds")
boost.varimp = add_rownames(boost.varimp)
boost.varimp <- mutate(boost.varimp, rowname = revalue(rowname, 
                                                       c("norm.log" = "log.pred",
                                         "norm.rf" = "rf.pred",
                                         "norm.boost" = "boost.pred")))
# order levels
boost.varimp$rowname <- factor(boost.varimp$rowname,
                               levels= boost.varimp$rowname)

@@

auc.list <- readRDS("blend_auc2.rds")
means = lapply(auc.list, mean) %>% unlist
greater = means[means> means["orig.log"]] %>% names
auc.melt = melt(auc.list)
auc.melt$greater = ifelse(auc.melt$L1 %in% greater, TRUE, FALSE)

@@

```{r blend_vs_initial}
auc.150 = readRDS("blend_auc150.rds")
mice.totAUC <- readRDS("mice_df_auc.rds")
compare.blends <- append(auc.150, mice.totAUC)

@@

reject.plot <- readRDS('reject_plot.rds')

@@

### 2016-08-05 finalwriteup/trial_code_eq.rmd
just conclusion left now
can't remember what I changed -- some grammar

$Overall\ Profit = Revenue\ from\ Project\ Invoices\ - (Employee + Business\ Costs)$

$f(x) = \beta_{0} + \sum_{j = 1}^{p} X_{j} \beta_{j}$

$RSS(\beta) = \sum_{i = 1}^{N} (y_{i} - f(x_{i}))^2$

$RMSE_b = \sqrt{\frac{\sum_{t = 1}^{n} (\widehat{y}_t - \bar y)^2}{n}}$

$$
\begin{aligned}
RMSE_b = \textrm{baseline root mean squared error} \\
n &= \textrm{number of data points} \\
\widehat{y} &= \textrm{predicted values of the response variable} \\
\bar y &= \textrm{mean 'return per dollar' for all projects} \\
\end{aligned}
$$

@@

>>>>

$Relative\ Profit_t = \frac{Profit_t}{\sum_{t = 1}^{n}\left|Profit_t\right|}$

$$
\begin{aligned}
Profit_t &= \textrm{Profit from project}\ t\textrm{. Negative values indicate loss making projects}
\end{aligned}
$$

@@

to losses of \$`r 156100 %>% format(digits = 2)`

```{r}
finance.plot = readRDS("plot_finance_100_4.rds" )

finance.plot$method <- factor(finance.plot$method, 
                                    levels=c("average", "simp.log","simp.boost",
                                             "comp.boost","comp.rf", "FWLS","orig.log",
                                             "orig.rf","orig.boost"))

ggplot(finance.plot %>% filter(method == "average") %>% slice(43:63), 
       aes(x= threshold, y= profit.ratio, colour = method)) + 
        geom_hline(yintercept = 100, colour = "darkgray", linetype = 2) + 
        geom_hline(yintercept = 0, colour = "darkgray", linetype = 1) + 
        geom_line() +
        geom_point() +
        scale_y_continuous(breaks = seq(0,110, by = 10)) + 
        scale_x_continuous(breaks = seq(0,1, by = .1)) + 
        theme(legend.position = "none", 
              panel.background = element_rect(fill = 'grey95'),
              panel.grid.minor = element_blank()) +
        labs(x = "Probability Threshold", 
             y = "Ratio of Model Profit to Historical Profit (%)",
             title = "Example Profit Curve for Simple Average Blended Model")

```
